---
title: "Quality control report"
output:
  pdf_document:
    #keep_tex: true  # for debugging

params:
  sample_name: "Sample 1"
  qc_result: NULL
  qc_plot: NULL
    
date: "`r format(Sys.Date(), '%B %d, %Y')`"
---
## Table of content

The following function is a quality control for the processing of the spatial samples. 

1) Check NAs in the matrix file

2) For the raw counts file columns:
  a) the column names are checked to make sure these are probe barcodes 
  b) this is checked by looking at the length of column name characters 
  
3) For the row counts file rows:
  a) the row names are checked to make sure they are genes 
  b) the number of characters of genes should be bigger than the row dimensions
  c) this is assessed using a summary of the character frequencies 
  d) additionally an example of rows is given 
  
4) The dimensions are checked to make sure the genes are rows and columns are barcodes

5) For the post PCA checks:
  a) All the loading factors for PC1 should be positive or negative (shown as +1 or -1). 
     A frequency table is made to check the frequency of loading factors
  b) Additionally for the PC1 removed raw counts table the number of NAs and SD of rows and columns are checked. The SD should never be 0. 
  
6) The seurat image needs to be plotted using features (housekeeping genes):
  a) This is to check the image 
  b) As well as if the scale factor is correct (this would cause problems with scaling)
  c) this should be checked for images original high res converted to low res 
  d) multiple housekeeping genes are used in case one is missing


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```
\newpage
# Quality control report for sample:

The report is for the following sample. 

```{r,results = "asis"}
cat(params$sample_name)
```

Below you can find the report details:

## Matrix NAs
```{r, results = "asis"}
cat("\n\n",params$qc_result[[1]],"\n")
```

## Rawcount column names
- Should be barcodes

```{r, results = "asis"}
cat("\n","The name of the first three columns are:","\n",params$qc_result[[2]],"\n")

cat("\n",params$qc_result[[3]])

```
```{r}
params$qc_result[[4]]
```

## Rawcount row names
- Should be genes 

```{r, results = "asis"}
cat("The maximum number of characters in the genes should be greater",
    "than number of rows.","\n",
    "The following shows: max character length in rownames, number of rows:", params$qc_result[[5]],
    "\n\n","Below shows a summary of the row's character length:","\n")
```
```{r}
params$qc_result[[6]]
```
```{r, results = "asis"}

cat("\n",params$qc_result[[7]], "\n")
```

## Raw count dimension's
- Should be genes*barcodes

```{r, results = "asis"}
cat("\n",params$qc_result[[8]], "\n")
```


Raw count duplicates and NAs should be 0. Duplicates should cause a feature plot error 

```{r, results = "asis"}
cat("\n",params$qc_result[[9]], "\n")
cat("\n",params$qc_result[[10]], "\n")
```

## Sign PC1 removed
- Should be the same 
- Final PC1 removed raw count has: rows as barcode spots and columns as genes  

```{r, results = "asis"}
params$qc_result[[11]]
cat("\n",params$qc_result[[12]], "\n")
cat("\n",params$qc_result[[13]], "\n")
cat("\n",params$qc_result[[14]], "\n")
```

## Sum of correlation matrix PCs
- The sum of the correlation between gene activity scores of PCs should be the same as the number of barcodes - 1 (since PC1 is removed)
- This is because only the same PCs i.e PC2-PC2 have a correlation of 1 (since they are the same)
- The rest should be 0 or very near 0 sincs PCs are independant
```{r, results="asis"}
cat("\n sum of correlation matrix, number of barcodes: ",params$qc_result[[15]], "\n")

```
\newpage

## Spatial feature plot of housekeeping genes: GADPH,ACTB, B2M
- a feature/gene is plotted to check image exists and scale factor is correct 

If this is abnormal it is likely due to a discrepency between hi res image and function 
This is an indication to double check this in the processing function

```{r}

params$qc_plot

```